Catecholamines and Hypoglycemia

### 2024-07-31 Christine Chan/EnVision CF/catecholamines_analysis.qmd
started catecholamines analysis

---
title: "Title"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

@@

```{r data cleaning}
# Import data cleaned for Katie's analysis
df <- read.csv("./Christine Chan/EnVision CF/Data_Clean/nacfc_2024_analysis_data.csv",
  na.strings = ""
)
glucose <- paste0("timepoint_", c(0, 10, 30, 60, 90, 120, 150, 180), "_min")
insulin <- gsub("timepoint_", "insulin_", glucose)
# Remove those missing last 3 glucose timepoints
df <- df[rowSums(!is.na(df[, tail(glucose, 3)])) > 0, ]
```

@@

```{r}
#| label: tbl-1
#| tbl-cap: Participant characteristics (overall)
df %>%
  group_by(study_id) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(
    redcap_data_access_group, age_at_current_date:origin_race, cftr_mutation_1,
    cftr_mutation_2
  ) %>%
  tbl_summary(by = redcap_data_access_group, missing_text = "Missing") %>%
  add_overall()
```

@@

```{r}
#| label: tbl-70
#| tbl-cap: BG < 70 at Any Point
df %>%
  select(
    age_at_current_date:origin_race, cftr_mutation_1, cftr_mutation_2,
    all_of(glucose), Hypo70, all_of(insulin), hypoglycemia_symptoms_yesno,
    iAUC30gluc:iAUC180ins, homa_ir
  ) %>%
  tbl_summary(by = Hypo70, missing_text = "Missing") %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

@@

```{r mixed models 70}
# Long OGTTs
long_df <- df %>%
  pivot_longer(timepoint_0_min:timepoint_180_min,
    names_to = "Time", values_to = "Glucose",
    names_pattern = "timepoint_(.*)_min"
  )
# Time as a factor
long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180))
# Fit
mod <- lmer(Glucose ~ Time * Hypo70 + (1 | study_id), data = long_df)
means <- data.frame(emmeans(mod, ~ Hypo70 | Time, pbkrtest.limit = 1e4))
# Long insulins
insulin_long_df <- df %>%
  pivot_longer(all_of(insulin),
    names_to = "Time", values_to = "Insulin",
    names_pattern = "insulin_(.*)_min"
  )
# Time as a factor
insulin_long_df$Time <- factor(insulin_long_df$Time,
  levels = c(0, 10, 30, 60, 90, 120, 150, 180)
)
# Fit
ins_mod <- lmer(Insulin ~ Time * Hypo70 + (1 | study_id), data = insulin_long_df)
ins_means <- data.frame(emmeans(ins_mod, ~ Hypo70 | Time, pbkrtest.limit = 1e4))
# Combine
means <- means %>% select(-SE, -df)
means$Variable <- "Glucose"
ins_means <- ins_means %>% select(-SE, -df)
ins_means$Variable <- "Insulin"
plot_df <- data.frame(rbind(means, ins_means))
levels(plot_df$Hypo70) <- c("No Hypoglycemia", "Hypoglycemia")
```

@@

>>>>

---
title: "Catecholamines and Hypoglycemia"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
    theme:
      light: flatly
      dark: darkly
editor: source
---

@@

```{r data cleaning}
# Import cleaned data
df <- read.csv("./Christine Chan/EnVision CF/Data_Clean/analysis_dataset.csv",
  na.strings = ""
)
# For this analysis, we only want those with catecholamines 
df = df %>% group_by(study_id,Date) %>%
  mutate(catecholamines = any(!is.na(Epinephrine)|!is.na(Norepinephrine))) %>%
  filter(catecholamines) %>% ungroup()
```

@@

BMI percentiles were calculated using the `childsds` R package, assuming that male = 1 and female = 2 for the `Sex` variable. Those with age > 20 were considered age = 20 for the purposes of percentile calculations.

```{r}
#| label: tbl-1
#| tbl-cap: Participant characteristics (overall)
df %>%
  group_by(study_id) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(
    redcap_data_access_group, age_visit:ethnicity, cftr_mutation_1,
    cftr_mutation_2
  ) %>%
  tbl_summary(by = redcap_data_access_group, missing_text = "Missing") %>%
  add_overall()
```

@@

## Demographics

IA0005, IA0006, IA0017, IA0042, and IA0079 had catecholamines measured at multiple timepoints but these are treated as independent in the table below.

```{r}
#| label: tbl-70
#| tbl-cap: BG < 70 at Any Point
df %>%
  filter(Timepoint %in% c(0, 120, 150, 180)) %>%
  pivot_wider(
    names_from = Timepoint, values_from = c(Epinephrine, Norepinephrine),
    id_cols = c(study_id:cftr_mutation_2,Hypo60,Hypo70)
  ) %>%
  select(Hypo70,age_visit:cftr_mutation_2,Epinephrine_0:Norepinephrine_180) %>%
  tbl_summary(by = Hypo70, missing_text = "Missing") %>%
  add_overall() %>%
  add_p() %>%
  separate_p_footnotes()
```

@@

```{r mixed models 70}
# Time as a factor
long_df$Time <- factor(long_df$Time, levels = c(0, 10, 30, 60, 90, 120, 150, 180))
# Fit
mod <- lmer(Glucose ~ Time * Hypo70 + (1 | study_id), data = long_df)
means <- data.frame(emmeans(mod, ~ Hypo70 | Time, pbkrtest.limit = 1e4))
# Long insulins
insulin_long_df <- df %>%
  pivot_longer(all_of(insulin),
    names_to = "Time", values_to = "Insulin",
    names_pattern = "insulin_(.*)_min"
  )
# Time as a factor
insulin_long_df$Time <- factor(insulin_long_df$Time,
  levels = c(0, 10, 30, 60, 90, 120, 150, 180)
)
# Fit
ins_mod <- lmer(Insulin ~ Time * Hypo70 + (1 | study_id), data = insulin_long_df)
ins_means <- data.frame(emmeans(ins_mod, ~ Hypo70 | Time, pbkrtest.limit = 1e4))
# Combine
means <- means %>% select(-SE, -df)
means$Variable <- "Glucose"
ins_means <- ins_means %>% select(-SE, -df)
ins_means$Variable <- "Insulin"
plot_df <- data.frame(rbind(means, ins_means))
levels(plot_df$Hypo70) <- c("No Hypoglycemia", "Hypoglycemia")
```

@@

## Demographics
